---
title: "Entering the tidyverse 01-23-2023"
author: "Abigail Griffin"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
## Entering the tidyverse
January 23 2023   
AGG   

* **tidyverse**: collection of packages that share philosophy, grammar (or how the code is structured), and data structures. More intuitive  
* **operators**: symbols that tell R to perform different operations (between variables, functions, etc.)  
+ e.g. arithmetic operators: + , - , *, /, ^, ~  
+ assignment operators: <- (how we create all variables)  
+ logical operators: !, &, | (or)
+ relational operators: ==, != (does not equal), >, <, >= (greater than or equal to), <=
+ today we will learn about miscellaneous operators: %>% (forward pipe operators), %in%

### Tidyverse
* Today we will be installing the tidyverse packages. You only need to install packages once. First `install.packages("tidyverse)` then `library(tidyverse)`
* library(tidyverse) loads in packages
* **dplyr**: new(er) packages provides a set of tools for manipulating data sets.  
+ specifically written to be fast  
+ individual functions that correspond to common operations
* , (comma) = both of these conditions have to be met for this to be filtered
* & (ampersand) = multiple conditions can be met
* | (line) = *or*

* 5 core verbs:
+ `filter()`: 
+ `arrange()`:
+ `select()`:
+ `groupby()` and `summarize()`:
+ `mutate()`:
* **Built in data set**
```{r eval=FALSE}
library(tidyverse)
library(dplyr)
data(starwars)
glimpse(starwars)
class(starwars)
```
+ **Tibble**: modern take on data frames. Keeps great aspects of df's and drops frustrating ones (change variables)
* If we want to clean up data:  
+ Check for NA's
```{r eval=FALSE}
anyNA(starwars) # are there any NA's? returns a logical # is.na #complete.cases
starwarsClean<-starwars[complete.cases(starwars[, 1:10]),] # removes all rows with NAs in all rows, the first 10 columns # second comma says give me all columns back
# anyNA(starwarsClean[,1:10]) # we have completely removed all NA's in first 10 columns
```
* use `**filter()**`: picks/subsets observations (ROWS) by their values
```{r eval=FALSE}
filter(starwarsClean, gender=="masculine" & height < 180) # can replace & (ampersand) with a comma which means and - both of these conditions have to be met for this to be filtered
filter(starwarsClean, gender=="masculine" , height < 180 , height > 100) # multiple conditions for the same variable
filter(starwarsClean, gender=="masculine" | gender=="feminine") # gender is masculine or feminine
```
* now working with `%in%` operator, a matching operator. Similar to the `==` but you can compare vectors of different length (can't do with `==`)
```{r eval=FALSE}
a<-LETTERS[1:10]
length(a) # how many elements are in this vector

#output of %in% depends of first vector
# a %in% b
# b %in% a
```
* use `%in%` to subset
```{r eval=FALSE}
eyes<-filter(starwars, eye_color %in% c("blue", "brown")) # subsetting characters with these eye colors
view(eyes) # opens up table in another tab
eyes2<-filter(starwars, eye_color == "blue" | eye_color == "brown")
view(eyes2)
```
* `arrange()`: reorders rows
```{r eval=FALSE}
sw<-arrange(starwarsClean, by=height) # arranges the whole data frame by the variable you specify
```
* can also use helper function `desc()`
```{r eval=FALSE}
arrange(starwarsClean, by=desc(height)) # arrange in descending order
```
* put in additional conditions
```{r eval=FALSE}
tail(sw) # missing values are added to the end ??
```
* `select()`: chooses variables (COLUMNS) by their names
```{r eval=FALSE}
select(starwarsClean, 1:10)
select(starwarsClean)
select(starwarsClean, name: species) # ??
select(starwarsClean, -(films:starships))
starwarsClean[,1:11]
```
* `everything()`: rearrange columns. a helper function that is useful if you have a couple variables that you want to move to the beginning
```{r eval=FALSE}
select(starwarsClean, name, gender, species, everything()) # reorders this but doesn't keep it reordered.
```
* `contains()`: another helper function
```{r eval=FALSE}
select(starwarsClean, contains ("color")) # others include: ends_with(), starts with(), num_range()
```
* `select()` can also rename columns
```{r eval=FALSE}
select(starwarsClean, haircolor=hair_color) # returns only renamed column
rename(starwarsClean, haircolor=hair_color) # returns whole data set
```
* `mutate()`: creates new variables using functions of existing variables
* let's create a new column that is height divided by mass
```{r eval=FALSE}
mutate(starwarsClean, ratio=height/mass)
starwars_lbs<-mutate(starwarsClean, mass_lbs=mass*2.2)
starwars_lbs<-select(starwars_lbs, 1:3, mass_lbs, everything()) # everything() is everything else. brought it to the front using select
glimpse(starwars_lbs)
```
* `transmute()`:
```{r eval=FALSE}
transmute(starwarsClean, mass_lbs=mass*2.2) # only returns mutated columns
transmute(starwarsClean, mass, mass_lbs=mass*2.2, height)
```
* `group_by()` and `summarize()`
```{r eval=FALSE}
summarize(starwarsClean, meanHeight=mean(height)) # throws NA if any NAs are in df - need to use na.rm
summarize(starwarsClean, meanHeight=mean(height), TotalNumber=n())
```
* use `group_by()` for maximum usefulness
```{r eval=FALSE}
starwarsGenders<-group_by(starwars, gender)
head(starwarsGenders) # lets you view first 6 rows, can also do head(starwarsGender, 10) to see first 10. Can also use tail(starwarsGender) to see last 6

summarize(starwarsGenders, meanHeight=mean(height, na.rm=TRUE), TotalNumber=n()) # TotalNumber=n() counts how many belong to each group
```

#### Piping `%>%`
* **piping** is used to emphasize a sequence of actions
* allows you to pass an intermediate result onto the next function (uses output of one function as input of the next)
* avoid if you need to manipulate more than one object/variable at a time or if a variable is meaningful
* formatting: should always have a space before the `%>%` followed by a new line
```{r eval=FALSE}
starwarsClean %>%
  group_by(gender) %>%
  summarize(meanHeight=mean(height, na.rm = TRUE, TotalNumber=n()))
# na.rm=TRUE removes NAs
# this data frame acts as %>% input
 # of this function
# much cleaner with piping!
```
* `case_when()`: is useful for multiple if/ifelse statements
```{r eval=FALSE}
starwarsClean %>%
  mutate(sp=case_when(species=="Human"~"Human", TRUE~"Non-Human"))
# if species = Human, put Human. Tilda tells what to put. If species is not true, put Non-Human.
# uses condition, puts "Human" if True in sp column, puts "Non-Human" if it's FALSE
```