---
title: "2023-01-24 ggplot2"
author: "Abigail Griffin"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
## ggplot2
January 24 2023  
AGG  

Load libraries.
```{r}
library(ggplot2)
library(ggthemes)
library(patchwork)
```

### Template for ggplot code
```{r eval=FALSE}
# p1<- ggplot(data=<DATA>, mapping=aes(x=xVar, y=yVar)) +
# <GEOM FUNCTION> + # geom_boxplot()
# if you wanted to map on a boxplot with your data
# print(p1)
```

### Load in a built-in data set
```{r}
d<-mpg
str(mpg)
library(dplyr)
glimpse(d)
```
* How to do qplot() (quick plots): for quick plotting, similar to what is in R base package. 
  + e.g. make a **histogram**  
```{r}
qplot(x=d$hwy)
qplot(x=d$hwy, fill=I("darkblue"), color=I("black"))
# need to put I before color 
```
* Make a **scatter plot**
```{r eval=FALSE}
library(dplyr)
library(ggplot2)
qplot(x=d$displ, y=d$hwy, geom=c("smooth", "point"), method="lm") # method makes it a linear distribution and not just following points
```
* Make it a **box plot**
```{r eval=FALSE}
qplot(x=d$fl, y=d$cty, geom="boxplot", fill=I("forestgreen"))
```
* Make a **basic bar plot**
```{r eval=FALSE}
qplot(x=d$fl, geom="bar", fill=I("forestgreen"))
# don't need y axis for bar plot so skip it
```
### Create some data with specified counts
```{r eval=FALSE}
x_trt<-c("Control", "Low", "High")
y_resp<-c(12, 2.5, 22.9)
qplot(x=x_trt, y=y_resp, geom="col", fill=I(c("forestgreen", "slategray", "goldenrod"))) # will print in alphabetical order
# uses vectors instead of whole data frame
```
* [To find colors click here](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf)
  + google "viridis color palette"

### ggplot: uses dataframes instead of vectors
```{r}
p1<-ggplot(data=d, mapping=aes(x=displ, y=cty, color=cyl)) +
  geom_point()
p1
# aes is very important: construct aesthetics mappings
p1 + theme_base() # changes to what it was when it was base R
p1 + theme_bw()
p1 + theme_classic() # removes background
p1 + theme_linedraw()
p1 + theme_dark() # adds dark background
p1 + theme_minimal() # no axis lines
p1 + theme_economist()
# google ggplot themes to find more
```
* [Find more ggplot themes here](https://ggplot2.tidyverse.org/reference/ggtheme.html)
* Can do further modifications
```{r}
p1 + theme_bw(base_size=30, base_family="serif")
p2<-ggplot(data=d, aes(x=fl, fill=fl)) +
  geom_bar()
p2

p2 + coord_flip() #flips x and y axes
p2 + coord_flip() + theme_classic(base_size=15, base_family="sans")
```
### More theme modifications
```{r}
p3<-ggplot(data=d, aes(x=displ, y=cty)) +
  geom_point(size=7, shape=21, color="magenta", fill="black") +
  xlab("Count") + # add x axis label
  ylab("Fuel") +
  labs(title="My Title Here", subtitle="My subtitle goes here") # can also do x=/y= for axis
p3

p3 + xlim(1,10) + ylim(0,35) # x is between 1 and 10, y is between 0 and 35
```
* Make another **boxplot**
```{r}
library(viridis)
cols<-viridis(7, option = "magma") # takes 7 hex color codes from the viridis color packages. # also plasma, turbo, viridis. # have to give it the same number of colors as you have groups
ggplot(data=d, aes(x=class,y=hwy, fill=class)) +
  geom_boxplot() +
  scale_fill_manual(values=cols)
```

### Stacking graphs
```{r eval=FALSE}
library(patchwork)
p1+p2+p3 # puts them all next to each other
p1/p2/p3 # puts them on top of each other
(p1+p2)/p3 # p1 and p2 on top over p3
```

```{r eval=FALSE}

```

```{r eval=FALSE}

```