---
title: "2023-01-26 Loading in Data"
author: "Abigail Griffin"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
## Loading in Data
January 26 2023   
AGG

### 1. Create and save a dataset:
* `write.table(x=varName, file="outputFileName.csv", header=TRUE, sep=",")`

### 2. These functions read in a data set
* `read.table()` and `read.csv()` are main two operators of how to read in a data set
  + `read.table(file="path/to/data.csv", header=TRUE, sep = ","`
      + usually better
  + `read.csv(file="data.csv", header = TRUE, sep = ",")`
* csv is best format for writing data into R

### Use RDS Object only when working in R: saveRDs
* RDS creates an R object that is loaded in R which you can send to people. It is helpful with large datasets - creates much smaller files to send
* `saveRDS(my_data, file="FileName.RDS")`: uses a variable you've created and named in R and saves it as its own object
* `readRDS("FileName.RDS")`
* `p<-readRDS("FileName.RDS")`

### Long v. Wide data formats
* **Long**: more rows than columns. Contains values that DO repeat in ID column
* **Wide**: more columns than rows. Contains values that do not repeat in the ID column
* e.g.
```{r}
library(tidyverse)
library(ggthemes)


head(billboard)
# this is wide format because there are so many columns
glimpse(billboard)

b1<-billboard %>%
  pivot_longer (
    cols = starts_with ("wk"), #cols = : specify which columns you want to make longer - put them in a row
  names_to = "Week", # name of new column which will contain the header names
  values_to = "Rank", # name of new column which will contain the values
  values_drop_na = TRUE) # excludes any rows which have NA
  ## only use view(b1) in R script or in console. Rmarkdown does not like this. Remove if knitting
```

* **pivot_wider**: best for making occupancy matrix
```{r}
glimpse(fish_encounters)
view(fish_encounters)

fish_encounters %>%
  pivot_wider(names_from = station, values_from=seen, values_fill=0) # names_from= which column you want to turn into multiple columns # values_from= which column contains the values for the new column cells
# now each fish is only represented once as a row and each station is represented if they are seen there. if never seen there is an NA. so add values_fill=0
# this is appropriate format if you want to create an occupancy matrix
```


### Publicly available data through Dryad
* a curated resource that makes research data discoverable, freely reusable, and citable.
* [Explore data sets here.](https://datadryad.org/search)
* Open dryad data using **`readtable()`**
```{r}
dryadData<-read.table(file="data/veysey-babbitt_data_amphibians.csv", header=TRUE, sep = ",") # header=TRUE because data do have a header # sep = "," because it's a csv file - separated by commas
glimpse(dryadData)
head(dryadData)
table(dryadData$species) # allows you to see different groups of character column
summary(dryadData$mean.hydro)

str(dryadData$species)
dryadData$species<-factor(dryadData$species, labels=c("Spotted Salamander", "Wood Frog")) # creating 'labels' to use for the plot
class(dryadData$treatment)

dryadData$treatment <- factor(dryadData$treatment, 
            levels=c("Reference",
                     "100m", "30m")) # Reference goes first, then 100m, then 30m

p<- ggplot(data=dryadData, 
       aes(x=interaction(wetland, treatment), # group treatment and wetland
           y=count.total.adults, fill=factor(year))) + geom_bar(position="dodge", stat="identity", color="black") +
  ylab("Number of breeding adults") +
  xlab("") +
  scale_y_continuous(breaks = c(0,100,200,300,400,500)) +
  scale_x_discrete(labels=c("30 (Ref)", "124 (Ref)", "141 (Ref)", "25 (100m)","39 (100m)","55 (100m)","129 (100m)", "7 (30m)","19 (30m)","20 (30m)","59 (30m)")) + #labels for x axis
  facet_wrap(~species, nrow=2, strip.position="right") +
  theme_few() + scale_fill_grey() + 
  theme(panel.background = element_rect(fill = 'gray94', color = 'black'), legend.position="top",  legend.title= element_blank(), axis.title.y = element_text(size=12, face="bold", colour = "black"), strip.text.y = element_text(size = 10, face="bold", colour = "black")) + 
  guides(fill=guide_legend(nrow=1,byrow=TRUE)) 

p

```
```{r warning=FALSE, message=FALSE}
# will not print out any text
```

