---
title: "Simple Data Analysis and More Complex Control Structures 01-30-2023"
author: "Abigail Griffin"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: console
---
## Simple Data Analysis and More Complex Control Structures
January 30 2023   
AGG

Loading in data from last week.
```{r}
dryadData<-read.table(file="data/veysey-babbitt_data_amphibians.csv", header=TRUE, sep = ",")
```

Set up libraries
```{r}
library(tidyverse)
library(ggthemes)
```

### Analyses
* For experimental designs, you usually have independent/explanatory variable (x-axis) vs. dependent/response variable (y-axis)
* Continuous v. discrete variables affect this
  + **Continuous variables**: range of numbers (e.g. height, weight, temperature)
  + **Discrete/Categorical variables**: categories (e.g. species, treatments, site)

### Basic linear regression analysis (2 continuous variables)
* Is there a relationship between the mean pool hydroperiod and the number of breeding frogs caught?
```{r}
glimpse(dryadData)
regModel<-lm(count.total.adults ~ mean.hydro, data=dryadData) # note y~x
summary(regModel) # use summary instead of print to get more pertinent stat data
hist(regModel$residuals)
summary(regModel)$"r.squared"
summary(regModel)[["r.squared"]]
# View(summary(regModel)) # View must be uppercase, comment out View because RMarkdown does not like

regPlot<-ggplot(data=dryadData, aes(x=mean.hydro, y=count.total.adults+1)) + # adding one = log transform?
  geom_point(size=2) +
  stat_smooth(method="lm", se=0.99) # sets standard error bars, default is 96
regPlot + theme_few()
```


### Basic ANOVA
* Was there a statistically significant difference in the number of adults among wetlands? Remember `y~x`
```{r}
ANOModel<-aov(count.total.adults~wetland, data=dryadData)
summary(ANOModel)
```

* Use `factor()` to tell ggplot these are different groups - otherwise just treats it as different numbers
```{r}
ANOModel<-aov(count.total.adults~factor(wetland), data=dryadData)
```

```{r}
dryadData %>%
  group_by(wetland) %>%
  summarize(avgHydro = mean(count.total.adults, na.rm=T), N=n())
```

* Going to use a **box plot** with an ANOVA - go hand in hand with visualizing these results
```{r}
dryadData$wetland
# class(dryadData$wetland) = integer. Need to override using factor
dryadData$wetland <- factor(dryadData$wetland)
class(dryadData$wetland)
ANOplot<- ggplot(data=dryadData, mapping=aes(x=wetland, y=count.total.adults)) +
  geom_boxplot()
ANOplot
```

* If we wanted to see what it looked like between species?
```{r}
ANOplot2<- ggplot(data=dryadData, mapping=aes(x=wetland, y=count.total.adults, fill=species)) + # mapping fills in the aesthetics, allows you to give aesthetics arguments 
  geom_boxplot() +
  scale_fill_grey() # use gear icon -> Chunk Output in Console to see plot in Console
ANOplot2
```

* This code saves pdf to repo folder: `ggsave(file="SpeciesBoxplots.pdf", plot=ANOplot2, device="pdf")`

### Logistic Regression
1. Simulate data set
```{r}
# xVar<-sort(rgamma())
```

* **gamma probabilities**: continuous variables that are always positive and have a skewed distribution
```{r}
xVar <- sort(rgamma(n=200,shape=5,scale=5))
yVar <- sample(rep(c(1,0),each=100),prob=seq_len(200)) # prob=seq_len gives it equal probability of being sampled bc n=200, same as if we did 1:200
logRegData <- data.frame(xVar,yVar) 
```

### Logistic Regression Analysis
```{r}
logRegModel<-glm(yVar~xVar, data=logRegData, family=binomial(link=logit)) # generic linear model, very important function if you're stats-oriented # latter part - how we specify we want a logistic regression
summary(logRegModel)$coefficients
logRegPlot<-ggplot(data=logRegData, aes(x=xVar, y=yVar)) +
  geom_point() +
  stat_smooth(method=glm, method.args=list(family=binomial))
logRegPlot # binomial is what tells R it is a logistic regression because we have 0 or 1

```

### Contingency table (chi-squared) Analysis
* Are there differences in the counts of males and females between species?
```{r}
countData<-dryadData %>%
  group_by(species) %>%
  summarize(Males=sum(No.males, na.rm=T), Females = sum(No.females, na.rm=T)) %>%
  select(-species) %>%
  as.matrix()
countData # gets data into a table

row.names(countData) = c("SS", "WF") # change row names
countData
```

* **Chi-squared**
  + Get residuals
```{r}
testResults<-chisq.test(countData)
testResults$residuals ##??????
testResults$expected ##????????
```
 
* **Mosaic plot (base R)** ??? base R
```{r}
mosaicplot(x=countData, col=c("goldenrod", "grey"), shade=FALSE)
```

* **Bar plot**
```{r}
countDataLong <- countData %>%
  as_tibble() %>%  # transform into tibble, doesn't like that it's a matrix
  mutate(Species = c(rownames(countData))) %>% # give me species data ???
  pivot_longer(cols=Males:Females, # take both columns and make them longer - put them in a row
  names_to = "Sex", # name of new column which will contain Males/Females
  values_to = "Count") # name of new column which contain the values for M/F
```

* **Plot bar graph**
```{r}
ggplot(data=countDataLong, mapping=aes(x=Species, y=Count, fill=Sex)) +
  geom_bar(stat="identity", position="dodge") + #plots bars next to each other
  scale_fill_manual(values=c("darkslategrey", "midnightblue"))
```

### Control Structures
* `if` and `ifelse` statements
  + `if` statement
`if (condition) {expression 1}`  
if it's THIS do THIS
  + `if (condition) {expression 1} else {expression 2}`  
  if it's THIS, do THIS, but if it's something else, do THAT
  + `if (condition1) {expression 1} else`
  + `if (condition2) {expression 2} else`
* else must appear on the same line as the expression
* use it for single values
```{r}
z<-signif(runif(1), digits=2) # give me one random uniform value rounded to second signif value
z>0.5 # this turns it into a logical

# use {} or not
if (z>0.8) {cat(z,"is a large number", "\n")} else # {cat} is concatinate fn # if z>0.8, return that z is a large number and put a line break ("/n")
  if (z<0.2) {cat(z, "is a small number", "\n")} else
  {cat(z, "is a number of typical size", "\n")
    cat("z^2=", z^2, "\n")
    y<-TRUE} # if its anywhere in between, say it's a number of typical size, then square it # y<-TRUE - can make up expression: if something's true do this
```

* ifelse to fill vectors
* ifelse(condition, yes, no)
* e.g. insect population where females lay 10.2 eggs on average, follows **Poisson distribution** (discrete probability distribution showing the likely number of times an event will occur). 35% parasitism where no eggs are laid. Let's make a distribution for 1000 individuals
```{r}
tester<-runif(1000)
eggs<-ifelse(tester>0.35, rpois(n=1000, lambda = 10.2),0) # if tester is > .35, make a distribution, if not assign a 0
hist(eggs) # make a histogram of eggs data
```

* e.g. create a vector of pvalues from a simulation and we want to create a vector to highlight the significant ones for plotting
```{r}
pVals<-runif(1000)
z<-ifelse(pVals<=0.025, "lowerTail", "nonSig") # if lower than/eq to this value say "lowerTail", if not say "nonSig" # using if else
z[pVals >= 0.975] <- "upperTail" # if gr than/eq to assign "upperTail" # use subsetting
table(z) # give tables w # of lowerTail/nonSig/upperTail values

```

### For Loops
* `for` **loops**: workhorse function for doing repetitive tasks. Universal in all computer languages. Controversial in R because:
  + often not necessary (we have vectorization in R)
  + very slow with certain operations (binding rows/columns together)
  + family of apply functions (pretty much do the same as for loops)
* Anatomy of `for` **loop**
  + `for(variable in sequence){starts loop`
  `body of for loop`
  `end of loop}`
* `var` is a counter variable that holds the current value of the loop (i, j, k)
  + the sequence is an integer vector that defines the start/end of the loop
```{r}
for (i in 1:5){
  cat("stuck in a loop", i, "\n")
  cat(3+2, "\n") # has nothing to do with which iteration you're in, will always produce the same value (5) in the loop
  cat(3+i, "\n") # i is whatever iteration/number it is in the sequence # in other words cat(3+loop #, line break)
}
print(i) # will print the last number
```

* Use a **counter variable** that maps to the position of each element. This is a **single for loop**.
```{r}
my_dogs<-c("chow", "akita", "malamute", "husky", "samoyed")
for(i in 1:length(my_dogs)){
  cat("i=", i, "my_dogs[i]=", my_dogs[i], "\n") # i value will be changing - in first iteration it'll be chow, then akita, malamute, etc # think of i like iteration number # end with line break bc good coding practice
}
```

* Can also create **double for loops**
```{r}
m<-matrix(round(runif(20), digits=2), nrow=5)
m
```

* Use a for loop to change these values - if we were just doing loop over **ROWS** (not double loop)
```{r}
for(i in 1:nrow(m)){
  m[i,] <- m[i,] + i # all columns, loop through one row at a time
} # this goes to first row, adds 1 to it, then to second row/iteration and adds 2 to it.
m
```

* Can do the same thing for the **COLUMNS** - loop over columns
  + *Note*: it's good practice to use `i` for columns and `j` for rows
```{r}
m<-matrix(round(runif(20), digits=2), nrow=5)
for(j in 1:ncol(m)){
  m[,j] <- m[,j] + j
}
m
```


* Loop over rows and columns
```{r}
m<-matrix(round(runif(20), digits=2), nrow=5)
for(i in 1:nrow(m)){
  for(j in 1:ncol(m)){
    m[i,j] <- m[i,j] + i + j # adds i and j to each cell, depends on location of cell
  } # end of column j loop
} # end of column i loop
print(m)
```

