---
title: "Assignment 9"
author: "Abigail Griffin"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
## Daily Assignment 9

### `For` Loops and `If` Statements

1. Using a `for` loop and an `if` statement, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a vector of integers with some zero values. Then, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output.
```{r}
vec<-(round(runif(n=50, min=0, max=1)))
vecfn<-function(vector){
  counter<-0
  for(i in 1:length(vector)){
  if (vector[i]==0) {counter<-counter+1}
  }
  return(counter)
}
vecfn(vec)

```


2. Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.
```{r}

intmate<-function(int1, int2){ #argument of fn is a placeholder for what the fn is going to take
  mate<-matrix(nrow=int1, ncol=int2)
  for(i in 1:nrow(mate)){
    for(j in 1:ncol(mate)){
      mate[i,j] <- i*j
    } # end of column j loop
  }
  return(mate)
}

intmate(3,4)

```


3. Use a for loop to repeat your previous functions from Weekly Assignment 2 (which simulate a data frame and extract a p-value) 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for the p-values (a total of 2 columns and 100 rows)
```{r eval=FALSE}
df_func<-function(means=c(mean1,mean2),sds=c(sd1,sd2)){
Type<-rep(c("Control", "Treatment"), each=25)
Value<-c(rnorm(n=25, mean=mean1, sd=sd1),(rnorm(n=25, mean=mean2, sd=sd2)))
typevalue_df<-data.frame(Type, Value, stringsAsFactors = FALSE)
return(typevalue_df)
}
q1df<-df_func()
ANOVA_Test<-aov(Value ~ Type, data=q1df)
pvalue_func<-function(data=NULL){
  p<-summary(ANOVA_Test)[[1]][["Pr(>F)"]][1]
  return(p)
}
for (i in 1:100){
  print(df_func)
  cat(3+2, "\n") # has nothing to do with which iteration you're in, will always produce the same value (5) in the loop
  cat(3+i, "\n")}

```



