---
title: "Assignment 9"
author: "Abigail Griffin"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
## Daily Assignment 9

### `For` Loops and `If` Statements

1. Using a `for` loop and an `if` statement, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a vector of integers with some zero values. Then, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output.
```{r}
vec<-(round(runif(n=50, min=0, max=1))) # create a vector of 50 rounded random uniform numbers between 0 and 1 
print(vec)

vecfn<-function(vector){
  counter<-0 # assign 0 to counter
  for(i in 1:length(vector)){ # if the value of the 'ith' iteration is 0, add 1 to counter
  if (vector[i]==0) {counter<-counter+1}
  }
  return(counter) # print the counter, which counted how many zeros were present
}
vecfn(vec) # run function with my vector

```


2. Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.
```{r}

intmate<-function(int1, int2){ # argument of fn is the placeholder for what the fn is going to take, this fn will take (int1, int2)
  mate<-matrix(nrow=int1, ncol=int2) # create a matrix named "mate" where the number of rows is = int1 and columns = int2
  for(i in 1:nrow(mate)){
    for(j in 1:ncol(mate)){
      mate[i,j] <- i*j # for the ith row and jth column of my matrix "mate", multiply the row number * column number. in other words, the value of [i,j] is i*j.
    } # end of column j loop
  }
  return(mate) # print this matrix
}
# Let's try it out:
intmate(3,4)

```


3. Use a for loop to repeat your previous functions from Weekly Assignment 2 (which simulate a data frame and extract a p-value) 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for the p-values (a total of 2 columns and 100 rows)
```{r}
df_func<-function(means=c(10,45),sds=c(1.5,2)){
Type<-rep(c("Control", "Treatment"), each=25) # create first column with 25 lines of both treatment and control
Value<-c(rnorm(n=25, mean=means[1], sd=sds[1]),(rnorm(n=25, mean=means[2], sd=sds[2]))) # create second column with random values which also fit mean/sd criteria
typevalue_df<-data.frame(Type, Value, stringsAsFactors = FALSE) # create df with "Type" and "Value" columns
return(typevalue_df) # return df please
}

df_func()
q1df<-df_func()

ANOVA_Test<-aov(Value ~ Type, data=q1df)
pvalue_func<-function(data=NULL){
  p<-summary(ANOVA_Test)[[1]][["Pr(>F)"]][1]
  return(p)
}
pval<-pvalue_func(data=ANOVA_Test) # assign the name pval to the pvalue returned from my ANOVA_Test

newdf<-data.frame(x=1:100, y=NA) # create a new df where the first row is the replicate number and the second row is NA's which will be filled with p values following the for loop.

for (i in 1:100){
  q1df<-df_func() # run this function df_func() and assign it to the name 'q1df'
  ANOVA_Test<-aov(Value ~ Type, data=q1df) # run an ANOVA based on my q1df data
  pval<-pvalue_func(data=ANOVA_Test) # run the function that finds the pvalue of the ANOVA data and assign this to 'pval'
  newdf[i,2] <- pval # in my newdf, place pval in the ith row, 2nd column (which previously was filled with NA's)
}

print(newdf)

```



