---
title: "Assignment3"
author: "Abigail Griffin"
date: "`r Sys.Date()`"
output: html_document
---
### This is question 1:
#### a.  
```{r}
x<-1.1
a<-2.2
b<-3.3
square<-function(z=NULL){
  z<-(x^a^b)
  print(z)
}
square(z=x)
```


#### b.  
```{r}
x<-1.1
a<-2.2
b<-3.3
square<-function(z=NULL){
  z<-(x^a)^b
  print(z)
}
square(z=x)
```

#### c.  
```{r}
x<-1.1
z<-3*(x^3)+2*(x^2)+1
print (z)
```

### This is question 2:
#### a: (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)   
```{r}
c((seq(from=1, to=8)), (seq(from=7, to=1)))
```
#### b: (1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)  
```{r}
rep(1:5, 1:5)
```
#### c: (5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)   
```{r}
rep(5:1, 1:5)
```

### This is question 3:
#### a: the serpent arrives and gets in the back of the line  
```{r}
queue <- c("sheep", "fox", "owl", "ant")
queue<-c(queue[1:4], "serpent")
```
#### b: the sheep enters the club (so disappears from the line)  
```{r}
queue<-c(queue[-1])
```

#### c: the donkey arrives and talks his way to the front of the line  
```{r}
queue<-c("donkey", queue[1:4])
```

#### d: the serpent gets impatient and leaves  
```{r}
queue<-c(queue[-5])
```

#### e: the aphid arrives and the ant invites him to cut in line (hint: check out the append function)  
```{r}
queue<-append(queue, "aphid", after=3)
```

#### f: Finally, determine the position of the aphid in the line (hint: use the which function).  
```{r}
which(queue=="aphid")
```


