---
title: "Assignment3"
author: "Abigail Griffin"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
## Daily Assignment 3
### 1: Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the variable z and print the value stored in each z.  
#### a. `x^a^b`   
```{r}
x<-1.1
a<-2.2
b<-3.3
square<-function(z=NULL){
  z<-(x^a^b)
  print(z)
}
square(z=x)
```

#### b. (x^a^)^b^     
```{r}
x<-1.1
a<-2.2
b<-3.3
square<-function(z=NULL){
  z<-(x^a)^b
  print(z)
}
square(z=x)
```

#### c. 3x^3^+2x^2^+1      
```{r}
x<-1.1
z<-3*(x^3)+2*(x^2)+1
print (z)
```

### 2: Using the rep and seq functions, create the following vectors:  
#### a: (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)   
```{r}
c((seq(from=1, to=8)), (seq(from=7, to=1))) # combine these values into a list
```
#### b: (1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)  
```{r}
rep(1:5, 1:5) # repeat 1-5 in increasing increments (+1) until 5
```
#### c: (5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)   
```{r}
rep(5:1, 1:5) 
```

### 3: Create a vector using the code: `queue <- c("sheep", "fox", "owl", "ant")`, where queue represents the animals that are lined up to enter an exclusive club, with the sheep at the front of the line. Using R expressions, update `queue` as:  
#### a: the serpent arrives and gets in the back of the line  
```{r}
queue <- c("sheep", "fox", "owl", "ant")
queue<-c(queue[1:4], "serpent")
print (queue)
```
#### b: the sheep enters the club (so disappears from the line)  
```{r}
queue<-c(queue[-1]) # remove the first element of the queue
print(queue)
```

#### c: the donkey arrives and talks his way to the front of the line  
```{r}
queue<-c("donkey", queue[1:4]) # add this element to the first position of my queue
print(queue)
```

#### d: the serpent gets impatient and leaves  
```{r}
queue<-c(queue[-5])
print(queue)
```

#### e: the aphid arrives and the ant invites him to cut in line (hint: check out the append function)  
```{r}
queue<-append(queue, "aphid", after=3) # append aphid immediately after position 3
print(queue)
```

#### f: Finally, determine the position of the aphid in the line (hint: use the which function).  
```{r}
which(queue=="aphid") # asks which position is the aphid in my queue?
```


