---
title: "Assignment 5"
author: "Abigail Griffin"
date: "`r Sys.Date()`"
output: html_document
---
#### Question 2: Create a list with the following named elements:
###### `my_matrix`, which is a 4 x 4 matrix filled with random uniform values, my_logical which is a 100-element vector of TRUE or FALSE values. You can do this efficiently by setting up a vector of random values and then applying an inequality to it., my_letters, which is a 26-element vector of all the lower-case letters in random order.
```{r}
my_data<-runif(n=16, min=0, max=20)
my_matrix<-matrix(data=my_data,nrow=4, ncol=4)
print(my_matrix)

random_numbers<-runif(n=100, min=0, max=100)
my_logical<- random_numbers>50

my_letters<-c(letters[1:26]) ## is this still a list?
my_letters<-sample(my_letters)
typeof(my_letters)

assignment4_list<-list(my_matrix, my_logical, my_letters)

```

#### Then, complete the following steps:
###### a: create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
```{r}
new_list<-list(assignment4_list[[1]][2,2],assignment4_list[[2]][2],assignment4_list[[3]][2])

```

###### b: use the typeof() function to confirm the underlying data types of each component in this list
```{r}
typeof(assignment4_list[[1]]) # double
typeof(assignment4_list[[2]]) # logical
typeof(assignment4_list[[3]]) # character
```

###### c: i) combine the underlying elements from the new list into a single atomic vector with the c() function. ii) What is the data type of this vector?
```{r}
# i)
single_vector<-c(new_list[[1]],new_list[[2]],new_list[[3]])

# ii)
typeof(single_vector)
# it has been coerced to character
```

#### Question 3: Create an empty data frame with two variables (= columns) and 26 observations (= rows) below:
   + call the first column/variable `my_unifs` and fill it with 26 random uniform values from 0 to 10
   + call the second variable `my_letters` and fill it with 26 capital letters in random order.
   + for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.
   + for the first variable, write a single line of R code to identify which rows have the missing values.
   + re-order the entire data frame to arrange the second variable in alphabetical order.
   + calculate the column mean for the first variable.
   + rename the second column ‘my_LETTERS’
```{r}
my_unifs<-runif(n=26, min=0, max=10)
my_letters<-sample(letters[1:26])
toupper(my_letters)

q3df<-data.frame(my_unifs, my_letters, stringsAsFactors = FALSE)
print(q3df)

set.seed(1)
q3df[sample(nrow(q3df), 4),1]<-NA
print(q3df)

q3df[is.na(q3df$my_unifs),]

abc_q3df<-q3df[order(q3df$my_letters), ]

mean(abc_q3df$my_unifs, na.rm=TRUE)

colnames(abc_q3df)[colnames(abc_q3df) == "my_letters"] ="my_LETTERS"
print(abc_q3df)


```


