---
title: "Assignment 4"
author: "Abigail Griffin"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
## Daily Assignment 4

### 1: Assign to the variable n_dims a single random integer between 3 and 10
```{r}
set.seed(1)
n_dims<-sample(3:10,1)
print(n_dims)
```

#### a: Create a vector of consecutive integers from 1 to n_dims^2^.
```{r}
vector<-seq(1:n_dims^2)
print(vector)
```
#### b: Use the sample function to randomly reshuffle these values.
```{r}
randomvector <- sample(vector)
print(randomvector)
```
#### c: create a square matrix with these elements. Print out the matrix.
```{r}
m<-matrix(data=randomvector, ncol=n_dims)
print(m)
```

#### d: find a function in r to transpose the matrix. Transpose flips elements in the matrix. Print it out again and note how it has changed.
```{r}
t(m)
```
#### e: calculate the sum and the mean of the elements in the first row and then the last row.
```{r}
mdataframe<-data.frame(m)
print(mdataframe)

rowSums(mdataframe[1,]) 
rowSums(mdataframe[n_dims,])

rowMeans(mdataframe[1,])
rowMeans(mdataframe[n_dims,])
```
#### f: set your code up so you can re-run it to create a matrix of a different size by only changing the n_dims value